Re-run, Repeat, Reproduce, Re-use, Replicate:\\Transforming Code into Scientific Contributions

### 2017-05-31 main.tex
Update on Overleaf.

\documentclass[a4paper,11pt]{article}

@@

@@

%Replication is a cornerstone of science. 
Reproducibility is a cornerstone of science. 
% Since science aims to discover general principles of how the world functions, 
% * <nicolas.rougier@inria.fr> 2017-04-24T15:49:43.902Z:
% 
% > Since science aims to discover general principles of how the world functions, 
% 
% 
% C'est un peu une déclaration à l'emporte pièce et je soupçconne qu'il y a une littérature énorme sur la définition de la science. Et à moins de vouloir s'aliéner toute l'astophysioque et la cosmologie, on pourrait au minimum étendre à l'univers.
% 
% ^ <fabien.benureau+overleaf@gmail.com> 2017-04-25T13:29:01.309Z:
% 
% Reopened by editing the source!
% From [wikipedia](https://en.wikipedia.org/wiki/World): "In a philosophical context, the world is the whole of the physical Universe, or an ontological world."
% 
% ^ <nicolas.rougier@inria.fr> 2017-04-25T13:45:04.565Z:
%
% Mouais.
%
% ^.
If an experimental result cannot be obtained more than once,
% it might be because of a fluke, a bias, a bug, a false positive or even a hack.
%it becomes, at best, nothing more than food for thought for future research. 
it merely becomes, at best, an observation that
may feed future research. 
Replication issues have received increased attention in recent years,
in the domain of medicine and psychology in particular.
One could think that computer science would mostly be shielded from such issues.
But precisely because it is easy to believe that
if a program runs once and gives expected results it will do so forever, 
crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \citep{Collberg:2016}. 
Computer science is plagued by replication problems,
in part, precisely because it seems impervious to them.

@@

The code is running and producing the expected results. 
The next step is to make sure that you can produce the same output over successive runs of your program, should you want to. 
In other words, the next step is to make your program deterministic,  producing {\em repeatable} output. 
Such repeatability is useful. 
If a run of the program produces a particularly puzzling results,  it allows you to scrutinize any step of the execution of the program by re-running it again. 
It is also the easiest way to prove that the program did indeed produce the published results.
Such repeatability is not always possible or easy \cite{Diethelm:2012}.
But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\

@@

Having made a software re-usable by others should offer a decent protection against bugs. Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
This is why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.
%Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.

In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
The reason is to be found in a differences in the respective implementations of the RNG between Python and Numpy.
As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
and has a period of 2**19937-1.
The underlying implementation in C is both fast and threadsafe.}
Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
Here, we want to make the Numpy RNG to match the Python RNG behvior.

@@

Using a very simple random walk example using Python, we've seen a variety of pitfalls of writing scientific code. The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be unacceptable in some circumstances. $R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard. This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published. But it's hardly the case today.\\

>>>>

% !TeX program = pdflatex
\documentclass[a4paper,11pt]{article}

@@

%\usepackage{mathspec}
%\setmainfont{Minion Pro Cond}
%\setmathsfont(Digits,Greek,Latin)[Numbers={Proportional}]{Minion Pro}
%\setmathrm{Minion Pro}
%\setsansfont{MyriadPro-Cond}

@@

%Replication is a cornerstone of science. 
Reproducibility is a cornerstone of science. 
% Since science aims to discover general principles of how the world functions, 
% * <nicolas.rougier@inria.fr> 2017-04-24T15:49:43.902Z:
% 
% > Since science aims to discover general principles of how the world functions, 
% 
% 
% C'est un peu une déclaration à l'emporte pièce et je soupçconne qu'il y a une littérature énorme sur la définition de la science. Et à moins de vouloir s'aliéner toute l'astophysioque et la cosmologie, on pourrait au minimum étendre à l'univers.
% 
% ^ <fabien.benureau+overleaf@gmail.com> 2017-04-25T13:29:01.309Z:
% 
% Reopened by editing the source!
% From [wikipedia](https://en.wikipedia.org/wiki/World): "In a philosophical context, the world is the whole of the physical Universe, or an ontological world."
% 
% ^ <nicolas.rougier@inria.fr> 2017-04-25T13:45:04.565Z:
%
% Mouais.
%
% ^.
If an experimental result cannot be obtained more than once,
% it might be because of a fluke, a bias, a bug, a false positive or even a hack.
%it becomes, at best, nothing more than food for thought for future research. 
it merely becomes, at best, an observation that
may feed future research \cite{Mesirov:2010,osc:2015}. 
Replication issues have received increased attention in recent years,
in the domain of medicine and psychology in particular \cite{Iqbal:2016}.
One could think that computer science would mostly be shielded from such issues.
But precisely because it is easy to believe that
if a program runs once and gives expected results it will do so forever, 
crucial steps to transform working code into meaningful scientific contributions are rarely undertaken \citep{Sandve:2013,Schwab:2000}. 
Computer science is plagued by replication problems,
in part, precisely because it seems impervious to them.\\

@@

The code is running and producing the expected results. 
The next step is to make sure that you can produce the same output over successive runs of your program, should you want to. 
In other words, the next step is to make your program deterministic,  producing {\em repeatable} output. 
Such repeatability is useful. 
If a run of the program produces a particularly puzzling results,  it allows you to scrutinize any step of the execution of the program by re-running it again. 
It is also the easiest way to prove that the program did indeed produce the published results.
Such repeatability is not always possible or easy \citep{Diethelm:2012, Courtes:2015}.
But for sequential programs not depending on exotic hardware, it often comes down to controlling the initialization of the pseudo-random number generators (RNG).\\

@@

Having made a software re-usable by others should offer a decent protection against bugs.
Unfortunately, this is not always the case and some recent cases have demonstrated the tremendous impact a bug can have in Science \citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}.
This is why, as explained by Peng et al. \cite{Peng:2006}, {\em the replication of important findings by multiple independent investigators is fundamental to the accumulation of scientific evidence}.\\
% Replicability is the assumption that any article that does not provide the code source makes: that the description it provides of the algorithms is sufficiently precise and complete to re-obtain the results it presents.

In our case, the replication of the random walk brings some unexpected discrepancies between the original Python version and the replicated Numpy version.
The reason is to be found in a differences in the respective implementations of the RNG between Python and Numpy.
As explained in the \href{https://docs.python.org/3.6/library/random.html}{Python documentation}, {\em Python uses the Mersenne Twister as the core generator. It produces 53-bit precision floats
and has a period of 2**19937-1.
The underlying implementation in C is both fast and threadsafe.}
Numpy RNG is also based on the Mersenne Twister generator but there are differences in the way seed is interpreted when initializing the generator.
Fortunately, Both Numpy and Python offer access to the internal state and we can {\em fix} it to make the behavior of one RNG to match the other RNG.
Here, we want to make the Numpy RNG to match the Python RNG behavior.\\

Without such explicit replication effort, it would be very hard to discover this faulty behavior and this why the ReScience journal has been created.
This journal targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible.

@@

Through a very simple random walk example using Python, we've seen a variety of pitfalls of writing scientific code. The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^{4}$ form is much preferred but requires a fair amount of work that might be too much of a burden in some circumstances. $R^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard. This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published. But it's hardly the case today.\\
